Understanding Factors Contributing to Application Approval in Urban Planning Decisions (2016-2024)

### 2024-12-04 paper/paper.qmd
Final Revision

This paper investigates the factors influencing application approvals submitted to the City of Toronto's Committee of Adjustment, using a Bayesian logistic regression model. By analyzing predictors such as application type, submission year, and planning district, the study provides findings into the dynamics of urban planning decisions. The model captures temporal trends, spatial variability, and procedural differences, offering a structured understanding of approval outcomes. Visualizations, including coefficient plots and effect plots, complement the analysis, helping to interpret the relationships between predictors and approval likelihood.

@@

One key finding is the temporal trend in approval rates. The analysis exhibits that approval odds have increased steadily over time, peaking in 2018. This trend may reflect evolving planning policies, greater efficiency in committee processes, or a shift in zoning priorities. The data suggests that the Committee of Adjustment has become more permissive in granting approvals, which could indicate a growing alignment between applications and city planning objectives. This result highlights how administrative processes adapt over time to urban development needs.

@@

While this paper provides findings into the factors influencing application approvals, there exists several limitations. These weaknesses cover data-related issues, methodological constraints, and broader contextual considerations.

1. **Data Limitations**
The analysis relies solely on the dataset provided by the City of Toronto's Committee of Adjustment, which, while complete in its coverage of applications, lacks certain important variables. Key contextual factors, such as community objections, project complexity, or the presence of political or economic pressures, are not captured. For instance, the dataset does not provide information about why certain applications are refused or approved beyond the recorded decision, limiting the ability to analyze qualitative or situational influences.

2. **Methodological Constraints**
The Bayesian logistic regression model assumes that the relationships between predictors and the outcome variable are additive and linear on the log-odds scale. This assumption may oversimplify real-world interactions between variables, such as how specific planning districts may interact with application types or temporal trends. The model's relatively low accuracy for both the training and test datasets suggests underfitting, indicating that it fails to capture the full complexity of the data and may not adequately explain the variance in outcomes. This limitation could stem from insufficient predictors, unaccounted non-linear relationships, or interactions among variables. Furthermore, the model does not fully account for potential multicollinearity among predictors, which could distort the estimated coefficients. While priors were carefully chosen, their influence on the results might vary with a different prior specification, introducing some degree of subjectivity into the analysis.

>>>>

This paper investigates the factors influencing application approvals submitted to the City of Toronto's Committee of Adjustment, using a Bayesian logistic regression model. By analyzing predictors such as application type, submission year, and planning district, the study examines the dynamics of urban planning decisions. The model captures temporal trends, spatial variability, and procedural differences, offering a structured understanding of approval outcomes. Visualizations, including coefficient plots and predicted probability plots, complement the analysis, helping to interpret the relationships between predictors and approval likelihood.

@@

One key finding is the temporal trend in approval rates. The analysis shows that approval odds have increased steadily over time, peaking in 2018. This trend may reflect evolving planning policies, greater efficiency in committee processes, or a shift in zoning priorities. The data suggests that the Committee of Adjustment has become more permissive in granting approvals, which could indicate a growing alignment between applications and city planning objectives. This result demonstrates how administrative processes adapt over time to urban development needs.

@@

While this paper analyzes the factors influencing application approvals, several limitations remain. These weaknesses include data-related issues, methodological constraints, and broader contextual considerations.

1. **Data Limitations**
The analysis relies solely on the dataset provided by the City of Toronto's Committee of Adjustment, which, while complete in its coverage of applications, lacks certain important variables. Contextual factors, such as community objections, project complexity, or political and economic pressures, are not captured. For instance, the dataset does not provide information about why certain applications are refused or approved beyond the recorded decision, limiting the ability to analyze qualitative or situational influences.

2. **Methodological Constraints**
The Bayesian logistic regression model assumes that the relationships between predictors and the outcome variable are additive and linear on the log-odds scale. This assumption may oversimplify real-world interactions between variables, such as how specific planning districts may interact with application types or temporal trends. The model's relatively low accuracy for both the training and test datasets suggests underfitting, failing to capture the data complexity and may not adequately explain the variance in outcomes. This could stem from insufficient predictors, unaccounted non-linear relationships, or interactions among variables. Furthermore, the model does not fully account for potential multicollinearity among predictors, which could distort the estimated coefficients. While priors were carefully chosen, their influence on the results might vary with a different prior specification, introducing some degree of subjectivity into the analysis.
